---
title: "A gentle introduction to the anthromes R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A gentle introduction to the anthromes R package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette presents an example workflow for using the `anthromes` R package for analyzing long-term human populations and land use.

First load the `anthromes` package.
```{r setup}
# analysis packages
library(anthromes)
library(stars)
library(tidyverse)
# visualization packages
library(gganimate)
library(patchwork)
```

Although the package is able to work with global data, we'll focus just on a subset of the eastern Mediterranean for simplicity.

```{r}
bbox <- st_bbox(c(xmin = 29, xmax = 39, ymin = 30, ymax = 40), crs = 4326)
```

# Download HYDE 3.2 data

Read in the anthromes data as a stars object. stars in an R package for working with space-time cubes like the HYDE 3.2 data, which are spatial rasters representing multiple time steps.

```{r}
hyde <- get_hyde(dir = '..') %>%
  st_crop(bbox) %>%
  st_as_stars() %>%
  # change the names to something more readable
  setNames(c('crops', 'grazing', 'rice', 'pop', 'irrigation', 'urban'))
```

A stars object prints two pieces of information, the attribute data (which is essentially a tibble that can be manipulated via typical tidyverse functions) and dimension information (which records the spatial and temporal dimensions of the object).

```{r}
hyde
```

Also load in the HYDE 3.2 supporting grids: land area, potential vegetation, and potential villages.
```{r}
inputs <- get_hyde('inputs', dir = '..') %>%
  st_crop(bbox) %>%
  setNames(c('land_area', 'pot_veg', 'regions', 'iso', 'pot_vill'))
```

You can easily plot these objects in ggplot using geom_stars().

```{r}
ggplot() +
  facet_wrap(~time) +
  coord_quickmap() +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  labs(title = 'HYDE 3.2 cropland', x = 'Latitude', y = 'Longitude')
```

You can easily animate these data using gganimate. 
```{r}
ggplot() +
  geom_stars(data = filter(hyde, time %in% time_steps_centuries)) +
  coord_quickmap() +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  # use transition_states() from gganimate instead of facet_wrap 
  gganimate::transition_states(time) +
  labs(title = 'HYDE 3.2 land use', subtitle = 'Cropland at {closest_state}', x = 'Latitude', y = 'Longitude')
```

By default, geom_stars will only plot the first attribute. If you'd like to plot multiple attributes at a time, the easiest way is to convert the attributes to an extra dimension.
```{r}
ggplot() +
  geom_stars(data = merge(hyde[1:2,,,8:13])) +
  facet_grid(attributes~time) +
  coord_quickmap() +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  labs(title = 'HYDE 3.2 land use', x = 'Latitude', y = 'Longitude')
```


# Apply anthromes classification

```{r}
anthromes <- get_anthromes(hyde, inputs)
```

```{r}
ggplot() +
  geom_stars(data = anthromes[,,,8:13]) +
  facet_wrap(~time) +
  coord_quickmap() +
  scale_fill_manual(values = anthrome_colors(), drop = TRUE) +
  theme_bw() +
  labs(title = 'Anthromes-12k', x = 'Latitude', y = 'Longitude')
```

```{r}
ggplot() +
  geom_stars(data = filter(anthromes, time %in% time_steps_centuries)) +
  transition_states(time) +
  coord_quickmap() +
  scale_fill_manual(values = anthrome_colors(), drop = FALSE) +
  theme_bw() +
  labs(title = 'Anthromes-12k', subtitle = 'Anthromes at {closest_state}', x = 'Latitude', y = 'Longitude')
```

```{r}
knitr::knit_exit()
```

# Discrete Global Grids

Rasters are great, but a hexagonal discrete global grid system would keep the cell areas the same and better represent shapes over the Earth's surface.
```{r}
dgg <- read_sf('../analysis/dgg_land.gpkg')[st_as_sfc(bbox),, op = st_covered_by]
```

```{r, echo = FALSE}
ggplot(dgg) +
  geom_sf() +
  theme_bw()
```


Use the get_dgg function to extract the HYDE 3.2 data using the DGG hexagons. It uses exact_extractr under the hood, which prints a progress bar by default.

```{r}
crops <-  dgg_extract(hyde, dgg, 'crops')
```

```{r}
ggplot() +
  geom_stars(data = (crops[,,8:13])) +
  facet_wrap(~time) +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  labs(title = 'HYDE 3.2 cropland', subtitle = 'Discrete global grid system', x = 'Latitude', y = 'Longitude')
```

Let's compare it to the 5' raster
```{r}
a <- ggplot() +
  geom_stars(data = hyde[,,,13]) +
  coord_quickmap() +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  labs(title = 'HYDE 3.2 cropland', x = 'Latitude', y = 'Longitude')

b <- ggplot() +
  geom_stars(data = crops[,,13]) +
  scale_fill_viridis_c(na.value = NA, name = expression(km^2)) +
  theme_bw() +
  labs(title = '', subtitle = 'Discrete global grid system', x = 'Latitude', y = 'Longitude')

a + b
```

Map this function over the list of HYDE variable names.
```{r}
hyde_dgg <- names(hyde) %>%
  map(~dgg_extract(hyde, dgg, .)) %>%
  do.call(c, .)
```

And extract the fixed inputs as well.
```{r}
# fold this into dgg_extract too?
inputs_dgg <- exact_extract(as(merge(inputs), 'Raster'), dgg, c('sum', 'mode')) %>%
   dplyr::select(land_area = sum.land_area, 
                pot_veg =  mode.pot_veg, 
                pot_vill = mode.pot_vill) %>%
    #rename_with(str_remove, pattern = 'sum.') %>% 
    mutate(geometry = dgg$geom) %>%
    st_as_sf() %>% 
    st_as_stars() %>%
    mutate(trees = pot_veg <= 8)
```

Apply the anthromes classifier to each level
```{r}
anthromes_dgg <- get_anthromes(hyde_dgg, inputs_dgg)
```

```{r}
ggplot() +
  geom_stars(data = anthromes_dgg[,,8:13]) +
  facet_wrap(~time) +
  scale_fill_manual(values = anthrome_colors(), drop = TRUE) +
  theme_bw() +
  labs(title = 'Anthromes-12k DGG', x = 'Latitude', y = 'Longitude')
```

```{r}
a <- ggplot() +
  geom_stars(data = anthromes[,,,13]) +
  coord_quickmap() +
  scale_fill_manual(values = anthrome_colors(), drop = TRUE) +
  theme_bw() +
  labs(title = 'Anthromes-12k', x = 'Latitude', y = 'Longitude')

b <- ggplot() +
  geom_stars(data = anthromes_dgg[,,13]) +
  scale_fill_manual(values = anthrome_colors(), drop = TRUE) +
  theme_bw() +
  labs(title = 'Anthromes-12k DGG', x = 'Latitude', y = 'Longitude')

a / b
```

# Anthromes summary statistics
 Use the anthrome_summary() function to produce formatted summary tables of the percent land area of under each anthrome type for each time period.
```{r}
anthrome_summary(anthromes_dgg, inputs_dgg)
anthrome_summary(anthromes, inputs)
```
anthrome_summary() also has a "by" argument, which adds an additional grouping factor. The resulting summaries are now in percent of land area under each anthrome in each time period in each of the grouping variables.
```{r}
anthrome_summary(anthromes_dgg, 
                 mutate(inputs_dgg, pot_veg = as.factor(pot_veg)), 
                 by = pot_veg) %>%
  select(all_of(time_steps_centuries))
```
You can also use the get_hyde_pop() function to extract population estimates from HYDE.
```{r}
get_hyde_pop(hyde_dgg, inputs_dgg) %>%
  ggplot(aes(time, pop, group = 1)) +
  geom_point() +
  stat_summary(fun.y=sum, geom="line")
```


# Statistical modeling

As stars objects, its easy to do statistical analyses on the data.
```{r}
pc <- hyde_dgg %>%
  select(-used) %>%
  as_tibble() %>%
  select(-geometry, -time) %>%
  prcomp(scale. = TRUE)

pc
```

```{r}
knitr::knit_exit()
```
